---
title: "p8105_hw2_xm2276.Rmd"
author: "XIAO MA"
date: "10/8/2021"
output: github_document
---

```{r}
library(tidyverse)
library(ggplot2)
library(readxl)
```

## Problem 0
### import the data
```{r}
trash_df = read_excel("data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx")
```

## problem 1
cleaning data
reading the sheet
```{r}
trash_df = read_excel("data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx",
                      sheet = "Mr. Trash Wheel") %>%
                      janitor::clean_names() %>%
                      drop_na(dumpster) %>%
                      mutate(round(sports_balls, digits = 0))

Prec2018 = read_excel("data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx",
                      sheet = "2018 Precipitation",
                      range = "A2:B14") %>%
                      mutate(Year = "2018")
Prec2018

Prec2019 = read_excel("data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx",
                      sheet = "2019 Precipitation",
                       range = "A2:B14") %>%
                      mutate(Year = "2019")
Prec2019
Prec_20182019 = bind_rows(Prec2018,Prec2019) %>%
                mutate(month.name[Month]) %>%
  relocate(Year)

```

The **number of observations of trash_df** contains 454 dumpsters through 9 years.
There are 17 numbers of variables of trash_df
The**key variables** are `r ls(trash_df)`
There are 12 number of observations of Prec2018 
The 3 numbers of variables of Prec2018
The**key variables** is total_amount of prep
There are 12 number of observations of Prec2019
There are 3 numbers of variables of Prec2019
The**key variables** are `r ls(Prec2019)`

The total precipitation in 2018 is 70.33.
The median number of sports balls in a dumpster in 2019 is 3.55 inch

The total precipitation in 2018 is 70.33.
The median number of sports balls in a dumpster in 2019 is 4.35 inch




## Problem 2
```{r}
pols_month_df = read_csv("data/fivethirtyeight_datasets/pols-month.csv")

pols_month_df = pols_month_df %>%
  separate(mon,into = c("year","month","day")) %>%
  mutate(
    year = as.integer(year),
    month = as.integer(month),
    day = as.integer(day)) %>%
mutate(month = month.name[month]) %>%
mutate(year = as.character(year)) %>%
drop_na(prez_gop) %>%
mutate(president = ifelse(prez_gop == 1, "gop","dem")) %>%
select(-prez_dem,-prez_gop,-day)
pols_month_df
```


Secondly,
```{r}
snp_df = read_csv("data/fivethirtyeight_datasets/snp.csv") %>%
      separate(date, into = c("month","day","year"), sep = "/") %>%
      mutate(year = as.numeric(year) + 2000) %>%
      mutate(year = ifelse(year >= "2050", year - 100, year)) %>%
      mutate(year = as.character(year)) %>%
      mutate(month = as.numeric(month)) %>%
      arrange(year, month) %>%
      mutate(month = month.name[month]) %>%
      mutate(year, month) %>%
      select(-day)
      
```

Thirdly,
```{r}
unemploy_df = read_csv("data/fivethirtyeight_datasets/unemployment.csv") %>% 
  pivot_longer(Jan:Dec,
  names_to = "month",values_to = "unemployment") %>%
  mutate(month = match(month, month.abb)) %>%
  mutate(month = month.name[month]) %>%
  janitor::clean_names() %>%
  mutate(year = as.character(year))

#join the data set
pols_snp = left_join(pols_month_df, snp_df, by = c("year","month"))
pols_snp_unemploy = left_join(pols_snp, unemploy_df, by = c("year", "month"))

```
Write a short paragraph about these datasets. Explain briefly what each dataset contained, and describe the resulting dataset (e.g. give the dimension, range of years, and names of key variables).

**The dimension of observations** of pols_month_df is `r nrow(pols_month_df)` and `r ncol(pols_month_df)`
**The range of the years** of pols_month_df is `r range(as.numeric(pols_month_df$year))`
**key variables** are `ls(pols_month_df)`

For the second dataset 
**The dimension of observations** of snp_df is `r nrow(snp_df)` and `ncol(snp_df)`
**The range of the years** of snp_df is `r range(as.numeric(snp_df$year))`
**key variables** are `ls(snp_df)`

For the third dataset
**The dimension of observations** of unemploy_df is `r nrow(unemploy_df)` and `r ncol(unemploy_df)`
**The range of the years** of unemploy_df is `r range(as.numeric(unemploy_df$year))`
**key variables** are `ls(unemploy_df)`



##problem3
```{r}
pop_names = 
  read_csv("data/Popular_Baby_Names.csv") %>%
  janitor::clean_names() %>%
  mutate(gender = tolower(gender)) %>%
  mutate(ethnicity = gsub(" ", "_",ethnicity)) %>%
  mutate(ethnicity = tolower(ethnicity)) %>%
  mutate(childs_first_name = tolower(childs_first_name)) %>%
  distinct( )
pop_names


##olivia
olivia = 
  filter(pop_names, childs_first_name == "olivia", gender == "female") %>%
  select(rank, year_of_birth, ethnicity) %>%
  arrange(year_of_birth)

##male names

pop_male_names = 
  filter(pop_names,gender == "male", rank == 1 ) %>%
  select(rank, year_of_birth, ethnicity) %>%
  arrange(year_of_birth)

name_2016 = filter(pop_names,year_of_birth == "2016", gender == "male", ethnicity == "white_non_hispanic")

ggplot(name_2016) + geom_point(aes(rank,count),color = "green")
ggsave("2016names.pdf")
```














